---
title: 'Replication of Wolf et al. 2021 : "as is" version'
author: "Florent Bédécarrats"
editor: visual
format:
  html: 
    code-fold: true
execute:
  warning: false
  error: false
---

## Introduction

Wolf et al. "Code availability" section states: "Analysis code is available at <https://github.com/wolfch2/PA_matching>." As the the words composing this link address indicate, it refers to a code repository called "PA_matching" hosted on github, created by the user with the identification "wolfch2".

The corresponding github repository contains the following files:

```{r}
library(httr)
library(tidyverse)

req <- GET("https://api.github.com/repos/wolfch2/PA_matching/git/trees/master?recursive=1")
filelist <- unlist(lapply(content(req)$tree, "[", "path"), use.names = F)
cat(filelist, sep = "\n")
```

To assess the reproducibility of this series of code files, we create a copy ("clone" in Github) of this repository at: [https://github.com/fBedecarrats/PA_matching](https://github.com/fBedecarrats/PA_matching).

## Retrieve data from Google earth engine

Display the content of the first file.

::: {.callout-code collapse="true"}

```{r}
#| echo: false
#| warning: false
#| class-output: "sourceCode python"

cat(readLines("001 - dl_GEE.py"), sep = "\n")
```
:::


